---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

```

```{css}
body{
  font-family: arial;
  font-size: 18px;
  background-color: #EEEEEE;
  color: black;
}

h1{
  font-size: 40px;
  color: #773500;
  padding: 20px;
}

h2{
  font-size: 28px;
  color: #19AA54;
  padding: 15px;
}

img{
  display: block;
  margin: auto;
}
```

## The birds that we see

<!---You need to change the contents of this Rmd! Replace the R code and markdown give below in this template document.--->

```{r load data from web}
library(tidyverse)

logged_data <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTx4qzXUuABXzuWuV6l5El_SuSGssLGo5kLCyBQ4uVAHDgKR9rjXpteaypJ_-aknNHhEc-zbg9XfSMS/pub?output=csv")
```
```{r rename variables}
latest_data <- logged_data %>%
  rename(date = 2,
         weather = 3,
         post_code = 4,
         setting = 5,
         bird_species = 6,
         temperature = 7)

# split date into date and time
date_time_placeholder <- strsplit(as.character(latest_data$date), " ")

date_time_df <- do.call(rbind, date_time_placeholder) %>%
  data.frame()

latest_data$date <- date_time_df[1]
latest_data$time <- date_time_df[2]

# group blackbird (female and male), chaffinch (female and male), and sparrow (hedge and house)
latest_data$bird_species[latest_data$bird_species == "Sparrow (House)"] <- "Sparrow"
latest_data$bird_species[latest_data$bird_species == "Sparrow (Hedge)"] <- "Sparrow"
latest_data$bird_species[latest_data$bird_species == "Blackbird (female)"] <- "Blackbird"
latest_data$bird_species[latest_data$bird_species == "Blackbird (male)"] <- "Blackbird"
latest_data$bird_species[latest_data$bird_species == "Chaffinch (female)"] <- "Chaffinch"
latest_data$bird_species[latest_data$bird_species == "Chaffinch (male)"] <- "Chaffinch"
latest_data$bird_species[latest_data$bird_species == "Mallard Duck"] <- "Duck"
latest_data$bird_species[latest_data$bird_species == "Grey Duck"] <- "Duck"
latest_data$bird_species[latest_data$bird_species == "Rosella (Eastern)"] <- "Rosella"
```
```{r summary values}
# The average temperature across all records is:
mean_temp <- mean(latest_data$temperature)

# The record with the most amount of weather events at once was:
# Below code was inspired by Zach Bobbit, Statsology, 2024 (https://www.statology.org/r-count-occurrences-of-character-in-string/)
weather_most <- regmatches(latest_data$weather, gregexpr(",", latest_data$weather)) %>%
  lengths() %>%
  max()

weather_most_words <- latest_data$weather[which(lengths(regmatches(latest_data$weather, gregexpr(", ", latest_data$weather))) == weather_most)][1]

# The number of unique bird species spotted was:
unique_birds <- latest_data$bird_species %>%
  unique() %>%
  length()
```
Interesting findings:

* The **average temperature** from the bird sightings was `r mean_temp`.
* The **most weather phenomena observed** at once was `r weather_most + 1`, which included `r weather_most_words`.
* The number of **unique birds** sighted was `r unique_birds`.

Here is a bar plot which shows the different bird species sighted:

```{r fig.width=12, fig.height=4}
# Bar plot to show the different bird species seen
latest_data %>%
  ggplot() +
  geom_bar(aes(x=bird_species),
           fill="#88FF99") +
  labs(title = "Count of bird species",
       x = "Bird species",
       y = "Count")
```
```{r}
bird_image_url <- c("Sparrow" = "https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/Dunnock_istock.jpg",
                    "Pidgeon" = "https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/Rock-Pigeon_Andrew-Walmsley.jpg",
                    "Gull" = "https://gardenbirdsurvey.nz/wp-content/uploads/cache/2022/02/Black-Backed-Gull_Paddy-Kemner/285460450.jpg",
                    "Blackbird" = "https://gardenbirdsurvey.nz/wp-content/uploads/cache/2022/02/Blackbird_Male_Andrew-Walmsley/2895393885.jpg",
                    "Song Thrush" = "https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/Song-Thrush_Tom-Marshall.jpg",
                    "Myna" = " https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/Myna_ashley-bradford.jpg",
                    "Duck" = "https://www.nzbirdsonline.org.nz/assets/95597/1691027917-mallarddrake_d803620-1024web.jpg",
                    "P큰keko" = "https://www.nzbirdsonline.org.nz/assets/95597/1691019412-28087715254_40361f0f45_o.jpg",
                    "Goose" = "https://www.nzbirdsonline.org.nz/assets/95597/1691027668-branta-canadensis-maxima_0561.jpg",
                    "Starling" = "https://gardenbirdsurvey.nz/wp-content/uploads/cache/2022/02/Starling_Andrew-Walmsley/2383985721.jpg",
                    "Parrot" = "https://www.nzbirdsonline.org.nz/assets/95597/1691021702-sulphur-crested-20cockatoo-2cred-20hill-2c230316-2c-59webnz.jpg",
                    "T큰카" = "https://gardenbirdsurvey.nz/wp-content/uploads/cache/2022/02/Tui_istock/3229547564.jpg",
                    "Kerer큰" = "https://gardenbirdsurvey.nz/wp-content/uploads/cache/2022/02/kereru/999649717.jpg",
                    "Magpie" = "https://gardenbirdsurvey.nz/wp-content/uploads/cache/2022/02/magpie/1480257895.jpeg",
                    "Morepork" = "https://www.nzbirdsonline.org.nz/assets/95597/1691022045-morepork.jpg",
                    "Albatross" = "https://www.nzbirdsonline.org.nz/assets/95597/1691015452-c6ea5c1a8dfdaa06db826b097a4bb354.jpg",
                    "Bellbird" = "https://gardenbirdsurvey.nz/wp-content/uploads/cache/2022/02/Bellbird_Andrew-Walmsley-002/3369328103.jpg",
                    "Chaffinch" = "https://gardenbirdsurvey.nz/wp-content/uploads/cache/2022/02/Chaffinch_Male_Jonothan-Harrod/2317722682.jpg",
                    "Fantail" = "https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/Fantail_istock.jpg",
                    "Goldfinch" = "https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/Goldfinch_Tom-Marshall.jpg",
                    "Greenfinch" = "https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/Greenfinch-Female_istock.jpg",
                    "Grey Warbler" = "https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/Grey-Warbler_-Satyaki-Hosmane.jpg",
                    "Heron" = "https://www.doc.govt.nz/thumbs/hero/contentassets/40c9692610874062b2ba71e542a2419f/white-heron-1920-3.jpg",
                    "Penguin" = "https://www.doc.govt.nz/thumbs/hero/contentassets/4d245bc7f1874034a2dfbc26538ae7b5/fiordland-crested-penguin-a-walmsley-1920-3.jpg",
                    "Rosella" = "https://gardenbirdsurvey.nz/wp-content/uploads/cache/2022/02/Rosella_istock/2422901731.jpg",
                    "Silvereye" = "https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/silvereye-bernard-spragg-CC0.jpg",
                    "Swallow" = "https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/Welcome-Swallow_Tom-Marshall.jpg",
                    "Yellowhammer" = "https://gardenbirdsurvey.nz/wp-content/uploads/2022/02/Yellowhammer_Female_Jacqui-nz.jpg",
                    "Other" = "https://www.doc.govt.nz/thumbs/hero/contentassets/a450e32f0b824531858d566404c21884/southern-brown-kiwi-tokoeka-stewart-island-photo-credit-alina-thiebes1920.jpg")

bird_count <- table(latest_data$bird_species)
bird_popular <- names(bird_count)[bird_count == max(bird_count)][1]

if (bird_popular %in% names(bird_image_url) == FALSE){
  bird_popular <- "Other"
}
```
The *most sighted bird* was `r bird_popular`
![](`r bird_image_url[bird_popular]`)

Here is a bar plot showing the different temperatures recorded for the amount of birds seen:

```{r fig.width=12, fig.height=4}
# Bar plot to show the different temperatures observed
latest_data %>%
  ggplot() +
  geom_bar(aes(x = temperature,
               fill = bird_species)) +
  labs(title = "Number of birds seen at different temperatures",
       x = "Temperature (degrees Celsius)",
       y = "Number of birds seen") +
  scale_fill_brewer(palette = "BrBG")
```
